\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\setheader{Acknowledgments}

First and foremost, I would like to express my deepest gratitude to my supervisor, \textbf{Éric PIETTE}, for his invaluable guidance, academic support, and trust throughout this project. His expertise in General Game Playing and his insightful feedback were essential in navigating the architectural challenges of this work.

\vspace{0.5cm}

I would also like to express my sincere appreciation to my readers, \textbf{Benoît DUHOUX}, \textbf{Achille MORENVILLE}, and \textbf{Hélène VERHAEGHE}, for taking the time to review this dissertation, for their constructive criticism, and for their dedication to ensuring its academic and scientific rigor.

\vspace{0.5cm}

Computational resources have been provided by the Consortium des Équipements de Calcul Intensif (CÉCI), funded by the Fonds de la Recherche Scientifique de Belgique (F.R.S.-FNRS) under Grant No. 2.5020.11 and by the Walloon Region. Specifically, I would like to acknowledge the use of the \textit{Dragon2} cluster at UCLouvain to collect the experimental metrics for both the Tabletop Games Framework (TAG) and the Ludii system, as well as the support and dedication of the local system administrators.

\vspace{0.5cm}

I am also deeply grateful to Google Gemini which was instrumental in the translation, grammatical correction, and rephrasing of significant portions of this dissertation.

\vspace{0.5cm}

Finally, I would like to thank my family from the bottom of my heart. This degree and this work belong just as much to you as they do to me. Thank you for being my anchor during the storms, for believing in me when doubt crept in, and for supporting me unconditionally when the path became difficult. Facing obstacles, setbacks, and uncertainties, we share this exact same strength: the refusal to ever give up. Your resilience has been my fuel, and this achievement stands as the ultimate symbol of our common refusal to ever back down. Thank you for everything.

\begin{flushright}
{\makeatletter\itshape
    Jérôme \\
    LECHAT, June 2026
\makeatother}
\end{flushright}